---
title: "Item Profiling and Forecasting - Time Series Class"
author: "Brad Pope"
date: "April, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary of the Dataset :

This is the sales history for a single UPC for a non-perishable product sold in a single retail location.

Here are the start, end, frequency and cycle information:

```{r Import, include=FALSE, echo=FALSE}

library(readr)
library(xts)
library(astsa)
library(readr)

cleanitem <- read_delim("C:/1 Brad/Dropbox/Data Science/Independent Study/cleanitem.txt", 
                        "\t", escape_double = FALSE, col_types = cols(Date = col_date(format = "%m/%d/%Y")), 
                        trim_ws = TRUE)

include=TRUE
#Converting the data to an xts
item <- xts(cleanitem$`Total volume`, order.by=as.Date(cleanitem$`Date`, "%m/%d/%Y"))

#Converting the data to a time series
c.item <- ts(cleanitem, frequency=365, start=c(2016, 268)) 

```

```{r Description, echo=FALSE}
# View the start and end dates of the dataset
#start(c.item)
#end(c.item)
#frequency(c.item)
cycle(c.item) # includes start, end and cycle

```
## Exploration

Exploring the dataset (plotting data: scatterplot, histogram):
```{r FreqDuration, echo=FALSE}
plot(item)
hist(item)

```

## Decomposition
Decomposition stage (plotting)


```{r, Decompose, echo=FALSE}
c.itemdecomp <- decompose(c.item, "additive")
plot(c.itemdecomp)
```

## Regression
building model, evaluation and interpretation

```{r, Regression, echo=FALSE}

#removed the trend
d_item=diff(item)
acf2(d_item)
plot(d_item)


#removing the cyclical nature of the data
dd_item=diff(d_item,lag=7)
plot(dd_item)

#Plot the ACF and PACF
acf2(dd_item, max.lag=60)

AR = arima(item, order = c(1,1,1)) # I tried this with the seasonal 7 digit to no avail
print(AR)

```
## Seasonal ARIMA
Building ARMA model (SARIMA) with a forecast

```{r, SARIMA model, echo=FALSE}
#Model a Seasonal ARIMA
x=sarima(item,1,1,1,0,1,1,7)
x

#Forecast the seasonal ARIMA
sarima.for(item,35,1,1,1,0,1,1,7)

```
## Model Diagnostics

ar1, ma1 and smal all have significant values
```{r, Model Diagnostics, echo = FALSE}

x$ttable

```
